---
title: "biostatistical methods homework 5"
output: pdf_document
---


```{r message=FALSE}
library(knitr)
library(tidyverse)
library(faraway)
library(broom)
```

R dataset ‘state.x77’ from library(faraway) contains information on 50 states from 1970s collected by US Census Bureau. The goal is to predict ‘life expectancy’ using a combination of remaining variables.

```{r}
life_data = as.data.frame(state.x77) %>%
  janitor::clean_names()
```

1. Explore the dataset and generate appropriate descriptive statistics and relevant graphs

```{r}
mean_and_sd = function(x) {
  
  if (!is.numeric(x)) {
    stop("Argument x should be numeric")
  } else if (length(x) == 1) {
    stop("Cannot be computed for length 1 vectors")
  }
  
  mean_x = mean(x)
  sd_x = sd(x)
  tibble(
    mean = mean_x, 
    sd = sd_x
  )
}
```

```{r message=FALSE}
attach(life_data)
```

```{r}
par(mfrow = c(2, 4))
boxplot(population, main = 'population')
boxplot(income,main = 'income' )
boxplot(illiteracy, main = 'illiteracy')
boxplot(life_exp, main = 'life_exp')
boxplot(murder, main = 'murder')
boxplot(hs_grad, main = 'hs_grad')
boxplot(frost, main = 'frost')
boxplot(area, main = 'area')
```

Population
```{r}
summary(population)
```

Income
```{r}
summary(income)
```

Illiteracy
```{r}
summary(illiteracy)
```

Life Exp
```{r}
summary(life_exp)
```

Murder
```{r}
summary(murder)
```

HS Grad
```{r}
summary(hs_grad)
```

Frost
```{r}
summary(frost)
```

Area
```{r}
summary(area)
```


2. Use automatic procedures to find a ‘best subset’ of the full model. Present the results and comment on the following:

```{r}
life_data.fit <- lm(life_exp ~., data=life_data)
summary(life_data.fit)
```

Backward elimination

```{r}
step1<-update(life_data.fit, . ~ . -area)
summary(step1)

step2<-update(step1, . ~ . -illiteracy)
summary(step2)

step3<-update(step2, . ~ . -income)
summary(step3)

step4<-update(step3, . ~ . -population)
summary(step4)
```

Using backward elimination, the final model contains murder, hs_grad and frost.

Forward elimination

```{r}
fit1 <- lm(life_exp ~ population, data=life_data)
tidy(fit1)

fit2 <- lm(life_exp ~ income, data=life_data)
tidy(fit2)

fit3 <- lm(life_exp ~ illiteracy, data=life_data)
tidy(fit3)

fit4 <- lm(life_exp ~ murder, data=life_data)
tidy(fit4)

fit5 <- lm(life_exp ~ hs_grad, data=life_data)
tidy(fit5)

fit6 <- lm(life_exp ~ frost, data=life_data)
tidy(fit6)

fit7 <- lm(life_exp ~ area, data=life_data)
tidy(fit7)
```

```{r}
forward1<-lm(life_exp~murder, data=life_data)
tidy(forward1)
```

```{r}
fit1 <- update(forward1, . ~ . +population)
tidy(fit1)
fit2 <- update(forward1, . ~ . +income)
tidy(fit2)
fit3 <- update(forward1, . ~ . +illiteracy)
tidy(fit3)
fit4 <- update(forward1, . ~ . +hs_grad)
tidy(fit4)
fit5 <- update(forward1, . ~ . +frost)
tidy(fit5)
fit6 <- update(forward1, . ~ . +area)
tidy(fit6)
```

```{r}
forward2 <- update(forward1, . ~ . + hs_grad)
tidy(forward2)
```

```{r}
fit1 <- update(forward2, . ~ . +population)
tidy(fit1)
fit2 <- update(forward2, . ~ . +income)
tidy(fit2)
fit3 <- update(forward2, . ~ . +illiteracy)
tidy(fit3)
fit4 <- update(forward2, . ~ . +frost)
tidy(fit4)
fit5 <- update(forward2, . ~ . +area)
tidy(fit5)
```

```{r}
forward3 <- update(forward2, . ~ . + frost)
tidy(forward3)
```

```{r}
fit1 <- update(forward3, . ~ . +population)
tidy(fit1)
fit2 <- update(forward3, . ~ . +income)
tidy(fit2)
fit3 <- update(forward3, . ~ . +illiteracy)
tidy(fit3)
fit4 <- update(forward3, . ~ . +area)
tidy(fit4)
```

```{r}
forward4 <- update(forward3, . ~ . + population)
tidy(forward4)
```

